Sakura Blossom Prediction Model for Japan

### 2024-11-28 paper/paper.qmd
need to look at the detail of words and some appendix

---
title: "Sakura Blossom Prediction Model for Japan"
subtitle: "Forecasting Sakura Blossom Using Bayesian Spline Regression"
author: 
  - Shanjie Jiao
thanks: "Code and data are available at: https://github.com/Jie-jiao05/Sakura-Blossom-Prediction-Model."
date: today
date-format: long
abstract: "This study develops a Bayesian hierarchical linear regression model to predict sakura blooming dates across Japan by analyzing key predictors such as temperature, latitude, and longitude. The model demonstrates strong predictive accuracy, capturing over 95% of the variability in flowering dates and revealing clear regional and climatic patterns. The findings provide valuable insights into the effects of climate change on sakura phenology, aiding tourism planning and ecological conservation. Future improvements could incorporate dynamic modeling techniques, additional environmental variables, and broader datasets to enhance the model's precision and adaptability."
format: pdf
toc: true
number-sections: true
bibliography: references.bib

@@

library(tidyverse)
library(dplyr)
library(here)
library(readr)
library(lubridate)
library(rvest)
library(palmerpenguins)
library(maps)
library(ggplot2)
library(knitr)
library(arrow)
library(rstanarm)
library(knitr)
library(plotly)

@@

result part

The structure of this paper is as follows: Section @sec-data details the data sources and the methodologies employed, including data scraping and manipulation techniques. Section @sec-model outlines the development of prediction models, specifically Linear Regression and Bayesian Spline Models, which are further analyzed in Section @sec-result. In Section @sec-dis, the impact of global warming on the sakura blossom period, along with real-life implementation and limitations of the study, will be discussed, providing insights for further improvement.

@@

For performing the analysis, we utilized several R packages. Tidyverse[@citetidyverse], Dplyr[@citedplyr], Here[@citehere], Readr[@citereadr], Lubridate[@citelubridate], Vest[@citervest)] for data cleaning and scraping.

@@

## Measurement
Our dataset, sourced from Alex Cookson's work [@Alex2020sakura], integrates temperature data scraped from the Japan Meteorological Agency [@Japanclimate]. The merged dataset contains 5,387 observations, aggregating average temperatures for the corresponding regions and flowering months. @fig-sample It includes detailed records on flowering dates, full bloom dates, and geographic locations. By compiling flowering times and geographic information for sakura blossoms across Japan since 1953 to 2019, this dataset provides comprehensive foundational data for studying the timing patterns and potential influencing factors of sakura blossom flowering.

```{r}
#| label: fig-sample   # for citation
#| fig-cap: Sample of Modern Sakura Data  #name of picture
#| echo: false
#| warning: false
#| message: false
#|

@@

# Display the sample as a clean table
kable(sakura_sample, 
      col.names = c("ID", "Location", "Latitude", "Longitude","Year","Month",
                    "Flower Day", "Full Bloom Day", " Mean Temp"))
```

For historical sakura data, since the earliest data in this dataset can be traced back to 812, however the accuracy of temperature measurements in the early years is questionable, and data being recorded only in the Kyoto area, there may be some unavoidable bias. Therefore, when building the prediction model, we will only use modern sakura data for fitting, and historical data will only serve as a comparison to help us understand the historical situation.

@@

# Load required library
library(ggplot2)
library(patchwork)

# Generate the histogram for Flowering Day
flowering_day_plot <- ggplot(sakura_data, aes(x = flower_day_in_numeric)) +
  geom_histogram(binwidth = 5, fill = "pink", color = "white") +
  labs(
    title = "Distribution of Flowering Day",
    x = "Flowering Day (Numeric)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Generate the histogram for Full Bloom Day
full_bloom_day_plot <- ggplot(sakura_data, aes(x = full_bloom_day_in_numeric)) +
  geom_histogram(binwidth = 5, fill = "#D8BFD8", color = "white", alpha = 0.9) +
  labs(
    title = "Distribution of Full Bloom Day",
    x = "Full Bloom Day (Numeric)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

@@

```{r}
#| label: fig-map   # For citation in Quarto
#| tbl-cap: "Recorded Geographical Information"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center
#| 
# Subset world map to Japan
japan_map <- map_data("world") %>% 
  filter(region == "Japan")

@@

```{r}
#| label: fig-temp   # For citation in Quarto
#| tbl-cap: "Change in Average Temperature Over Years with Trend Line"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#### Data Preparation ####
# Calculate yearly average temperature
yearly_avg_temp <- tem_data %>%
  group_by(year) %>%
  summarize(mean_temp_c = mean(mean_temp_c, na.rm = TRUE), .groups = "drop")  # Handle missing values

# Generate the plot with less specific year axis and a fitted line
ggplot(yearly_avg_temp, aes(x = year, y = mean_temp_c)) +
  geom_line(color = "#0072B2", size = 1) +    # Line plot with blue color
  geom_point(color = "#CC79A7", size = 2) +   # Points with pink color
  geom_smooth(method = "lm", color = "#000000", linetype = "dashed", se = FALSE, size = 1) + # Linear fit
  scale_x_continuous(
    breaks = seq(min(yearly_avg_temp$year), max(yearly_avg_temp$year), by = 10), # Set ticks every 10 years
    labels = scales::number_format(accuracy = 1) # Ensure clean formatting for labels
  ) +
  labs(
    title = "Change in Average Temperature Over Years with Trend Line",
    x = "Year",
    y = "Mean Temperature (°C)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12)
  )

@@

### Latitude and Longitude with Temperature
@fig-cor demonstrates how temperatures vary across different locations based on their geographical coordinates. A positive relation could be observed, with lower temperatures observed at higher latitudes, such as in northern Japan, and gradually increasing temperatures as the coordinates approach regions closer to the equator.
```{r}
#| label: fig-cor   # For citation in Quarto
#| tbl-cap: "Temperature Variation with Latitude and Longitude"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center
#| 
# Create a 2D scatter plot with color indicating temperature
ggplot(sakura_data, aes(x = latitude, y = longitude, color = month_mean_temp)) +
  geom_point(size = 3) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(
    title = "Temperature Variation with Latitude and Longitude",
    x = "Latitude",
    y = "Longitude",
    color = "Mean Temp (°C)"
  ) +
  theme_minimal()
```

# Model

@@

## Model set-up
This study uses a Bayesian Hierarchical Linear Regression model to analyze the relationship between sakura flowering dates and various predictors, implemented using the stan_glmer function from the rstanarm package @rstanarm in R [@citeR] and use the default priors from `rstanarm`. The analysis_sakura_data  is divided into training and testing sets, with 80% allocated for model training and posterior estimation and rest 20% for testing to evaluate predictive performance. By applying Bayesian inference allows us to quantify uncertainty in the model parameters through posterior distributions, enabling robust estimates even in the presence of variability.

@@

In this model, random effects are included to account for group-level variability at cross regions and year levels. Region-specific random effects (\(\gamma_{\text{region}}\)) capture local environmental differences, such as microclimates or soil conditions, allowing the model to adjust predictions for region with consistently earlier or later flowering patterns. Similarly, year-specific random effects (\(\delta_{\text{year}}\)) account for deviations in flowering dates caused by year-to-year climatic anomalies, such as warmer winters or extreme weather events. These random effects are modeled as zero-centered normal distributions with variances (\(\sigma_{\gamma}\) and \(\sigma_{\delta}\)) that reflect the variability among regions and years. By incorporating random effects, the model accommodates unobserved heterogeneity, will improves prediction accuracy, and realistically captures the hierarchical structure of the data.

@@

```{r}
#| label: fig-view  # For citation in Quarto
#| tbl-cap: "Training and Testing Data Evaluation Results"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center

@@

# Result

@@

```{r}
#| label: fig-rela_result
#| fig-cap: "Relationships Between Predictors and Flowering Day"
#| echo: false
#| warning: false
#| message: false
# Load required libraries
library(ggplot2)
library(tidyr)

@@

```{r}
#| label: fig-pred
#| fig-cap: "Observed vs. Predicted Flowering Dates"
#| echo: false
#| warning: false
#| message: false
# === Step 3: Generate Predictions ===
# Training predictions
train_predictions <- posterior_predict(flower_model)
train$predicted_flower_day <- colMeans(train_predictions)

@@

ggplot(data = combined_data, aes(
  x = flower_day_in_numeric, 
  y = predicted_flower_day, 
  color = Set
)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Observed vs. Predicted Flowering Dates",
    x = "Observed Flowering Dates",
    y = "Predicted Flowering Dates",
    color = "Dataset"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )

@@

```{r}
#| label: tbl-rmse  # For citation in Quarto
#| tbl-cap: "Training and Testing Data Evaluation Results"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center
# === Step 3: Generate Predictions ===
# Training predictions
train_predictions <- posterior_predict(flower_model)
train$predicted_flower_day <- colMeans(train_predictions)

@@

# === Step 6: Display Evaluation Metrics Table ===
kable(
  metrics_table,
  digits = 3,
  caption = "Training and Testing Data Evaluation Results",
  booktabs = TRUE
)

@@

The @fig-fix highlights the relationships between key predictors and sakura flowering dates based on the Bayesian hierarchical model. The intercept (\(-47.4751\)) represents the expected flowering date when all predictors are zero, though it’s mainly a baseline without much practical interpretation since predictors like temperature and location are scaled.

Monthly mean temperature has a clear and significant effect. For every 1°C increase, the flowering date is delayed by about 0.33 days, with a credible interval (\([0.2599, 0.4017]\)) that excludes zero. Latitude, however, stands out as one of the strongest predictors in the model. With a coefficient of \(4.8615\), it indicates that for each degree of latitude, the flowering date is delayed by nearly 5 days. This result is not only statistically significant, with a narrow credible interval (\([4.4334, 5.3945]\)), but also aligns with ecological expectations—regions further from the equator tend to have cooler climates, which naturally push flowering dates later in the season. Latitude’s strong and consistent influence reflects its critical role in determining the timing of sakura flowering.

Longitude, in contrast, has a much smaller and less certain effect. The coefficient (\(-0.2695\)) suggests that flowering may occur earlier as longitude increases, but the credible interval (\([-0.7217, 0.1194]\)) includes zero. This uncertainty means longitude’s impact is less clear, and it likely plays a minor role compared to temperature and latitude.

```{r}
#| label: fig-fix   # For citation in Quarto
#| tbl-cap: "Fixed Effects Coefficients with 95% Credible Intervals"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center
# Extract fixed effects coefficients and standard errors
fixed_effects <- flower_model$coefficients[1:4]  # Extract only the first 4 fixed effects
fixed_se <- flower_model$ses[1:4]  # Extract standard errors for fixed effects

@@

# Display the table in a formatted manner
library(knitr)
kable(
  fixed_effects_table,
  digits = 4,
  caption = "Fixed Effects Coefficients with 95% Credible Intervals"
)

@@

The random effects for year and region play a critical role in capturing temporal and spatial variability in sakura flowering dates and are shown in @tbl-ram . The Root Mean Squared Error (RMSE) of **4.282** shows that, on average, flowering dates deviate by approximately 4.3 days due to differences between years and regions. This highlights the substantial adjustments made by the random effects to account for annual climate trends and regional environmental factors. Similarly, the Mean Absolute Error (MAE) of **3.186** indicates that typical adjustments are around 3.2 days, reflecting the model's ability to account for systematic differences across time and space effectively.

The variance explained by the random effects (\(R^2 = 1.000\)) demonstrates that the model perfectly partitions the variability attributed to year and region. Year-based random effects capture temporal patterns, such as global warming or specific climate anomalies, ensuring that flowering trends align with observed climate changes. Region-based random effects, on the other hand, account for spatial heterogeneity influenced by geographic and environmental differences, such as latitude and local climate conditions. Together, these random effects ensure that the model accurately reflects both temporal and spatial dynamics.

```{r}
#| label: tbl-ram   # For citation in Quarto
#| tbl-cap: "Performance Metrics for Random Effects"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center
# === Step 1: Extract Random Effects ===
# Extract station-level and year-level random effects
random_effects <- ranef(flower_model, condVar = TRUE)

@@

# Print the table
library(knitr)
kable(
  random_effects_metrics,
  digits = 3,
  caption = "Performance Metrics for Random Effects",
  booktabs = TRUE
)

@@

# Discussion

@@

@@

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

@@

# Appendix {-}

@@

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. 
This shows the density overlay plot demonstrates a strong alignment between the observed sakura flowering dates and the model's posterior predictions. The observed density, represented by the dark blue line, closely matches the predictive densities (lighter blue lines) across most of the distribution, indicating that the model captures the central tendency and major patterns effectively. The primary peak around 80 is well-represented, showing the model's strength in predicting the most common flowering dates. Additionally, the secondary peak around 120 aligns reasonably well, suggesting the model's ability to handle more complex patterns, such as bimodal distributions. However, slight discrepancies are observed in the tails of the distribution, particularly below 40 and above 140, where the predictive densities deviate from the observed data. These divergences indicate potential challenges in capturing extreme values or outliers, which may be addressed by including additional predictors, such as microclimatic conditions or species-specific traits.

In @fig-ppcheckandposteriorvsprior we compare the posterior with the prior. This shows...

@@

library(bayesplot)
library(gridExtra)

>>>>

---
title: "Sakura Blossom Prediction Model for Japan"
subtitle: "Forecasting Sakura Blossom Using Bayesian Hierarchical Regression"
author: 
  - Shanjie Jiao
thanks: "Code and data are available at: https://github.com/Jie-jiao05/Sakura-Blossom-Prediction-Model."
date: today
date-format: long
abstract: "This study develops a Bayesian hierarchical linear regression model to predict sakura blooming dates across Japan by analyzing key predictors such as temperature, latitude, and longitude. The model demonstrates strong predictive accuracy, capturing over 95% of the variability in flowering dates and revealing clear regional and climatic patterns. The findings provide valuable insights into the effects of climate change on sakura phenology, aiding tourism planning and ecological conservation. Future improvements could incorporate dynamic modeling techniques, additional environmental variables, and broader datasets to enhance the model's precision and adaptability."
format: pdf
toc: true
number-sections: true
bibliography: references.bib

@@

library(tidyverse)
library(dplyr)
library(here)
library(readr)
library(lubridate)
library(rvest)
library(palmerpenguins)
library(maps)
library(ggplot2)
library(knitr)
library(arrow)
library(rstanarm)
library(plotly)
library(patchwork)
library(tidyr)
library(bayesplot)
library(gridExtra)

@@

The Bayesian hierarchical model build in this research predicts sakura flowering dates with high accuracy, achieving RMSE values of 3.1 (training) and 3.4 (testing) while explaining over 95% of variability.  Notably, temperature and latitude emerge as significant predictors, revealing earlier blooming in southern regions and highlighting strong sensitivity to climatic and geographic factors.

The structure of this paper is as follows:  @sec-data details the data sources and the methodologies employed, including data scraping and manipulation techniques.  @sec-model outlines the development of prediction models, specifically Linear Regression and Bayesian Spline Models, which are further analyzed in  @sec-result. In  @sec-dis, the impact of global warming on the sakura blossom period, along with real-life implementation and limitations of the study, will be discussed, also how sakura flowering condition from history to present, and will providing insights for further improvement. Further model detail will be presented in @sec-model-details

@@

For performing the analysis, we utilized several R packages. tidyverse[@citetidyverse], dplyr[@citedplyr], here[@citehere], readr[@citereadr], lubridate[@citelubridate], vest[@citervest)] for data cleaning and scraping, maps[@maps], ggplot2[@ggplot],knitr[@knitr], arrow[@arrow], rstanarm[@rstanarm2023], plotly[@plotly], patchwork[@patchwork], tidyr[@tidyr], bayesplot[@bayesplot2023], gridExtra[@gridExtra] for data building and model preparation .

@@

## Measurement
Our dataset, sourced from Alex Cookson's work [@Alex2020sakura], integrates temperature data scraped from the Japan Meteorological Agency [@Japanclimate]. The merged dataset contains 5,387 observations, aggregating average temperatures for the corresponding regions and flowering months. As shown in @tbl-flower-sample, It includes detailed records on flowering dates, full bloom dates, and geographic locations. By compiling flowering times and geographic information for sakura blossoms across Japan since 1953 to 2019, this dataset provides comprehensive foundational data for studying the timing patterns and potential influencing factors of sakura blossom flowering.

```{r}
#| label: tbl-flower-sample
#| tbl-cap: Sample of Modern Sakura Data
#| echo: false
#| warning: false
#| message: false

@@

# Display the sample as a clean table
kable(sakura_sample, 
      col.names = c("ID", "Location", "Latitude", "Longitude", "Year", "Month",
                    "Flower Day", "Full Bloom Day", "Mean Temp"))
```
For historical sakura data, since the earliest data in this dataset can be traced back to 812, however the accuracy of temperature measurements in the early years is questionable, and data being recorded only in the Kyoto area, there may be some unavoidable bias. Therefore, when building the prediction model, we will only use modern sakura data for fitting, and historical data will only serve as a comparison to help us understand the historical situation.

@@

# Generate the histogram for Flowering Day
flowering_day_plot <- ggplot(sakura_data, aes(x = flower_day_in_numeric)) +
  geom_histogram(binwidth = 5, fill = "pink", color = "white") +
  labs(
    title = "Flowering Day",
    x = "Flowering Day (Numeric)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Generate the histogram for Full Bloom Day
full_bloom_day_plot <- ggplot(sakura_data, aes(x = full_bloom_day_in_numeric)) +
  geom_histogram(binwidth = 5, fill = "#D8BFD8", color = "white", alpha = 0.9) +
  labs(
    title = "Full Bloom Day",
    x = "Full Bloom Day (Numeric)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

@@

```{r}
#| label: fig-map   # For citation in Quarto
#| fig-cap: "Recorded Geographical Information"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center
#| 
# Subset world map to Japan
japan_map <- map_data("world") %>% 
  filter(region == "Japan")

@@

```{r}
#| label: fig-temp   # For citation in Quarto
#| fig-cap: "Change in Average Temperature Over Years with Trend Line"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#### Data Preparation ####
# Calculate yearly average temperature
yearly_avg_temp <- tem_data %>%
  group_by(year) %>%
  summarize(mean_temp_c = mean(mean_temp_c, na.rm = TRUE), .groups = "drop")  # Handle missing values

# Generate the plot with less specific year axis and a fitted line
ggplot(yearly_avg_temp, aes(x = year, y = mean_temp_c)) +
  geom_line(color = "#0072B2", size = 1) +    # Line plot with blue color
  geom_point(color = "#CC79A7", size = 2) +   # Points with pink color
  geom_smooth(method = "lm", color = "#000000", linetype = "dashed", se = FALSE, size = 1) + # Linear fit
  scale_x_continuous(
    breaks = seq(min(yearly_avg_temp$year), max(yearly_avg_temp$year), by = 10), # Set ticks every 10 years
    labels = scales::number_format(accuracy = 1) # Ensure clean formatting for labels
  ) +
  labs(
    title = "Average Temperature Change Over Years",
    x = "Year",
    y = "Mean Temperature (°C)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12)
  )

@@

### Latitude and Longitude with Temperature
@fig-cor demonstrates how temperatures vary across different locations based on their geographical coordinates. A positive relation could be observed, with lower temperatures observed at higher latitudes, such as in northern Japan, and gradually increasing temperatures as the coordinates approach regions closer to the equator.
```{r}
#| label: fig-cor   # For citation in Quarto
#| fig-cap: "Temperature Variation"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center
#| 
# Create a 2D scatter plot with color indicating temperature
ggplot(sakura_data, aes(x = latitude, y = longitude, color = month_mean_temp)) +
  geom_point(size = 3) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(
    title = "Temperature Variation with Latitude and Longitude",
    x = "Latitude",
    y = "Longitude",
    color = "Mean Temp (°C)"
  ) +
  theme_minimal()
```

# Model {#sec-model}

@@

## Model set-up
This study uses a Bayesian Hierarchical Linear Regression model to analyze the relationship between sakura flowering dates and various predictors, implemented using the stan_glmer function from the rstanarm package @rstanarm in R [@citeR] and use the default priors from rstanarm @rstanarm. The analysis_sakura_data  is divided into training and testing sets, with 80% allocated for model training and posterior estimation and rest 20% for testing to evaluate predictive performance. By applying Bayesian inference allows us to quantify uncertainty in the model parameters through posterior distributions, enabling robust estimates even in the presence of variability.

@@

In this model, random effects are included to account for group-level variability at cross regions and year levels. Region-specific random effects $\gamma_{\text{region}}$ capture local environmental differences, such as microclimates or soil conditions, allowing the model to adjust predictions for regions with consistently earlier or later flowering patterns. Similarly, year-specific random effects $\delta_{\text{year}}$ account for deviations in flowering dates caused by year-to-year climatic anomalies, such as warmer winters or extreme weather events. These random effects are modeled as zero-centered normal distributions with variances $\sigma_{\gamma}$ and $\sigma_{\delta}$ that reflect the variability among regions and years. By incorporating random effects, the model accommodates unobserved heterogeneity, improves prediction accuracy, and realistically captures the hierarchical structure of the data.

@@

```{r}
#| label: fig-view  # For citation in Quarto
#| fig-cap: "Training and Testing Data Evaluation Results"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center

@@

# Result {#sec-result}

@@

```{r}
#| label: fig-rela_result
#| fig-cap: "Predictors and Outcome Variable Relations"
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-pred
#| fig-cap: "Observed vs. Predicted Flowering Dates"
#| echo: false
#| warning: false
#| message: false

# === Step 3: Generate Predictions ===
# Training predictions
train_predictions <- posterior_predict(flower_model)
train$predicted_flower_day <- colMeans(train_predictions)

@@

ggplot(data = combined_data, aes(
  x = flower_day_in_numeric, 
  y = predicted_flower_day, 
  color = Set
)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Fit Between Observed and Predicted Flowering Dates",
    x = "Observed Flowering Dates",
    y = "Predicted Flowering Dates",
    color = "Dataset"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )

@@

```{r}
#| label: tbl-rmse  # For citation in Quarto
#| tbl-cap: "Training and Testing Data Evaluation Results"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center

# === Step 3: Generate Predictions ===
# Training predictions
train_predictions <- posterior_predict(flower_model)
train$predicted_flower_day <- colMeans(train_predictions)

@@

# === Step 6: Display Evaluation Metrics Table ===
kable(
  metrics_table,
  digits = 3,
  booktabs = TRUE
)

@@

The @fig-fix highlights the relationships between key predictors and sakura flowering dates based on the Bayesian hierarchical model. The intercept -47.4751 represents the expected flowering date when all predictors are zero, though it’s mainly a baseline without much practical interpretation since predictors like temperature and location are scaled.

Monthly mean temperature has a clear and significant effect. For every 1°C increase, the flowering date is delayed by about 0.33 days, with a credible interval [0.2599, 0.4017] that excludes zero. Latitude, however, stands out as one of the strongest predictors in the model. With a coefficient of 4.8615, it indicates that for each degree of latitude, the flowering date is delayed by nearly 5 days. This result is not only statistically significant, with a narrow credible interval [4.4334, 5.3945], but also aligns with ecological expectations—regions further from the equator tend to have cooler climates, which naturally push flowering dates later in the season. Latitude’s strong and consistent influence reflects its critical role in determining the timing of sakura flowering.

Longitude, in contrast, has a much smaller and less certain effect. The coefficient -0.2695 suggests that flowering may occur earlier as longitude increases, but the credible interval [-0.7217, 0.1194] includes zero. This uncertainty means longitude’s impact is less clear, and it likely plays a minor role compared to temperature and latitude.

```{r}
#| label: fig-fix   # For citation in Quarto
#| tbl-cap: "Fixed Effects Coefficients with 95% Credible Intervals"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center

# Extract fixed effects coefficients and standard errors
fixed_effects <- flower_model$coefficients[1:4]  # Extract only the first 4 fixed effects
fixed_se <- flower_model$ses[1:4]  # Extract standard errors for fixed effects

@@

# Display the table in a formatted manner
library(knitr)
kable(
  fixed_effects_table,
  digits = 4,
)

@@

The random effects for year and region play a critical role in capturing temporal and spatial variability in sakura flowering dates and are shown in @tbl-ram . The Root Mean Squared Error (RMSE) of 4.282 shows that, on average, flowering dates deviate by approximately 4.3 days due to differences between years and regions. This highlights the substantial adjustments made by the random effects to account for annual climate trends and regional environmental factors. Similarly, the Mean Absolute Error (MAE) of 3.186 indicates that typical adjustments are around 3.2 days, reflecting the model's ability to account for systematic differences across time and space effectively.

The variance explained by the random effects (R^2 = 1.000) demonstrates that the model perfectly partitions the variability attributed to year and region. Year-based random effects capture temporal patterns, such as global warming or specific climate anomalies, ensuring that flowering trends align with observed climate changes. Region-based random effects, on the other hand, account for spatial heterogeneity influenced by geographic and environmental differences, such as latitude and local climate conditions. Together, these random effects ensure that the model accurately reflects both temporal and spatial dynamics.

```{r}
#| label: tbl-ram   # For citation in Quarto
#| tbl-cap: "Performance Metrics for Random Effects"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center

# === Step 1: Extract Random Effects ===
# Extract station-level and year-level random effects
random_effects <- ranef(flower_model, condVar = TRUE)

@@

# Print the table
library(knitr)
kable(
  random_effects_metrics,
  digits = 3,
  booktabs = TRUE
)

@@

# Discussion {#sec-dis}

@@

## History and Modern Sakura

Since the history_data only documents the blooming status of cherry blossoms in Kyoto and the tem_data records March temperatures, in @fig-his, we analyze data from sakura_data and tem_data to focus on Kyoto's blooming trends and climate in March.

The left panel illustrates that historically, cherry blossoms in Kyoto typically bloomed between the 90th and 120th days of the year, corresponding to late March to early April. In contrast, modern flowering days have shifted earlier, often occurring before day 100.

Meanwhile, the right panel shows that for much of history, March temperatures fluctuated between 2.5°C and 10°C. However, in recent years, temperatures have risen sharply, now ranging from 5°C to 12°C. This rapid temperature increase aligns with the earlier blooming days depicted in the left panel, highlighting a clear and significant relationship between rising temperatures and cherry blossom blooming in Kyoto. This linkage underscores the impact of climate change on seasonal cycles and phenology.

```{r}
#| label: fig-his
#| fig-cap: "History and Modern"
#| echo: false
#| warning: false 
#| message: false
#| fig-align: center

# Process historical flowering data
historical_flowering <- history_data %>%
  select(year, flower_doy) %>%
  rename(flower_day = flower_doy)

# Filter modern flowering data for Kyoto
modern_flowering_kyoto <- sakura_data %>%
  filter(station_name == "Kyoto") %>%
  select(year, flower_day_in_numeric) %>%
  rename(flower_day = flower_day_in_numeric)

# Process historical temperature data for March
historical_march_temp <- history_data %>%
  select(year, temp_c_recon) %>%
  rename(temperature = temp_c_recon) %>%
  mutate(month = 3) # Assume March data

# Filter modern temperature data for Kyoto in March
modern_march_temp_kyoto <- tem_data %>%
  filter(station_name == "Kyoto", month == "Mar") %>%
  select(year, mean_temp_c) %>%
  rename(temperature = mean_temp_c)

# Create flowering plot with new colors and shortened title
flowering_plot <- ggplot() +
  geom_line(data = historical_flowering, aes(x = year, y = flower_day), 
            color = "#9370DB", size = 1.2, alpha = 0.8) + # Medium purple for historical data
  geom_line(data = modern_flowering_kyoto, aes(x = year, y = flower_day), 
            color = "#708090", size = 1.2, alpha = 0.8) + # Slate gray for modern data
  labs(
    title = "Flowering Days",
    x = "Year",
    y = "Flowering Day"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 11, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 11),
    legend.position = "none"
  )

# Create temperature plot with new colors and shortened title
temperature_plot <- ggplot() +
  geom_line(data = historical_march_temp, aes(x = year, y = temperature), 
            color = "#4682B4", size = 1.2, alpha = 0.8) + # Steel blue for historical data
  geom_line(data = modern_march_temp_kyoto, aes(x = year, y = temperature), 
            color = "#708090", size = 1.2, alpha = 0.8) + # Cadet blue for modern data
  labs(
    title = "Temperatures",
    x = "Year",
    y = "Temperature (°C)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 11,  face = "bold", hjust = 0.5),
    axis.title = element_text(size = 11),
    legend.position = "none"
  )

# Combine the two plots side by side with updated title
combined_plot <- flowering_plot + temperature_plot + 
  plot_annotation(title = "Flowering Days and Temperatures in Kyoto", 
                  theme = theme(plot.title = element_text(size = 13, face = "bold", hjust = 0.5)))

# Display the combined plot
print(combined_plot)

@@

@@

# Appendix {#sec-app}

@@

In @fig-ppcheckandposteriorvsprior we implement a posterior predictive check. 
This shows the density overlay plot demonstrates a strong alignment between the observed sakura flowering dates and the model's posterior predictions. The observed density, represented by the dark blue line, closely matches the predictive densities (lighter blue lines) across most of the distribution, indicating that the model captures the central tendency and major patterns effectively. The primary peak around 80 is well-represented, showing the model's strength in predicting the most common flowering dates. Additionally, the secondary peak around 120 aligns reasonably well, suggesting the model's ability to handle more complex patterns, such as bimodal distributions. However, slight discrepancies are observed in the tails of the distribution, particularly below 40 and above 140, where the predictive densities deviate from the observed data. These divergences indicate potential challenges in capturing extreme values or outliers, which may be addressed by including additional predictors, such as microclimatic conditions or species-specific traits.

@@
